The Journey from Wild to Text Book Data: A Case Study from the National Longitudinal Survey of Youth

### 2021-09-01 Report/report_alt.tex

\title{The Journey from Wild to Text Book Data: A Case Study of the National Longitudinal Survey of Youth}

@@

% Pandoc citation processing

@@

``Open data'' is data that are freely accessible, modifiable, and shareable by anyone for any purpose (Open Knowledge Foundation 2021). This type of data can be useful as example data in statistical text books and for research purposes. However, open data are often, what we might call ``wild data'', because it requires substantial cleaning and tidying to tame it into text book shape. Huebner, Vach, and Cessie (2016) emphasize that making the data cleaning process accountable and transparent is imperative. Documenting the data cleaning is essential for the integrity of downstream statistical analyses and model building (Huebner et al. 2020).

@@

Despite its importance, this IDA stage is often undervalued and neglected (Chatfield 1985). There are few research papers that document the data cleaning (Wickham 2014). Furthermore, the decisions made in this stage often go unreported in the sense that IDA is often performed in an unplanned and unstructured way, and is only shared among restricted parties (Huebner et al. 2020).

@@

\caption{\label{tab:untidy-data}The Untidy Form of the NLSY79 Raw Data}
\centering
\begin{tabular}[t]{r|r|r|r|r|r|r}
\hline
CASEID\_1979 & HRP1\_1979 & HRP2\_1979 & HRP3\_1979 & HRP4\_1979 & HRP5\_1979 & HRP1\_1980\\
\hline
\rowcolor{gray!6}  1 & 328 & NA & NA & NA & NA & NA\\
\hline
2 & 385 & NA & NA & NA & NA & 457\\
\hline
\rowcolor{gray!6}  3 & 365 & NA & 275 & NA & NA & 397\\
\hline
4 & NA & NA & NA & NA & NA & NA\\
\hline
\rowcolor{gray!6}  5 & 310 & 375 & NA & NA & NA & 333\\
\hline
6 & NA & NA & NA & 250 & NA & 275\\
\hline
\end{tabular}
\end{table}

Consequently, the data should be tidied and wrangled first to extract the demographic and employment variables that we want to put in the final data set. We mainly used \texttt{tidyr} (H. Wickham 2020b), \texttt{dplyr} (H. Wickham, François, et al. 2020), and \texttt{stringr} (H. Wickham 2019) to do this job.

@@

\CommentTok{## tidy the date of birth data}
\NormalTok{dob_tidy <-}\StringTok{ }\NormalTok{new_data_qnames }\OperatorTok{%>%}
\StringTok{  }\NormalTok{dplyr}\OperatorTok{::}\KeywordTok{select}\NormalTok{(CASEID_}\DecValTok{1979}\NormalTok{,}
         \KeywordTok{starts_with}\NormalTok{(}\StringTok{"Q1-3_A~"}\NormalTok{)) }\OperatorTok{%>%}
\StringTok{  }\KeywordTok{mutate}\NormalTok{(}\DataTypeTok{dob_year =} \KeywordTok{case_when}\NormalTok{(}
                    \CommentTok{# if the years recorded in both sets match, take 79 data}
                    \StringTok{`}\DataTypeTok{Q1-3_A~Y_1979}\StringTok{`} \OperatorTok{==}\StringTok{ `}\DataTypeTok{Q1-3_A~Y_1981}\StringTok{`} \OperatorTok{~}\StringTok{ `}\DataTypeTok{Q1-3_A~Y_1979}\StringTok{`}\NormalTok{,}
                    \CommentTok{# if the year in the 81 set is missing, take the 79 data}
                    \KeywordTok{is.na}\NormalTok{(}\StringTok{`}\DataTypeTok{Q1-3_A~Y_1981}\StringTok{`}\NormalTok{) }\OperatorTok{~}\StringTok{ `}\DataTypeTok{Q1-3_A~Y_1979}\StringTok{`}\NormalTok{,}
                    \CommentTok{# if the sets don't match for dob year, take the 79 data}
                    \StringTok{`}\DataTypeTok{Q1-3_A~Y_1979}\StringTok{`} \OperatorTok{!=}\StringTok{ `}\DataTypeTok{Q1-3_A~Y_1981}\StringTok{`} \OperatorTok{~}\StringTok{ `}\DataTypeTok{Q1-3_A~Y_1979}\StringTok{`}\NormalTok{),}
        \DataTypeTok{dob_month =} \KeywordTok{case_when}\NormalTok{(}
                    \CommentTok{# if months recorded in both sets match, take 79 data}
                    \StringTok{`}\DataTypeTok{Q1-3_A~M_1979}\StringTok{`} \OperatorTok{==}\StringTok{ `}\DataTypeTok{Q1-3_A~M_1981}\StringTok{`} \OperatorTok{~}\StringTok{ `}\DataTypeTok{Q1-3_A~M_1979}\StringTok{`}\NormalTok{,}
                    \CommentTok{# if month in 81 set is missing, take the 79 data}
                    \KeywordTok{is.na}\NormalTok{(}\StringTok{`}\DataTypeTok{Q1-3_A~M_1981}\StringTok{`}\NormalTok{) }\OperatorTok{~}\StringTok{ `}\DataTypeTok{Q1-3_A~M_1979}\StringTok{`}\NormalTok{,}
                    \CommentTok{# if sets don't match for dob month, take the 79 data}
                    \StringTok{`}\DataTypeTok{Q1-3_A~M_1979}\StringTok{`} \OperatorTok{!=}\StringTok{ `}\DataTypeTok{Q1-3_A~M_1981}\StringTok{`} \OperatorTok{~}\StringTok{ `}\DataTypeTok{Q1-3_A~M_1979}\StringTok{`}\NormalTok{),}
        \CommentTok{# flag if there is a conflict between dob recorded in 79 and 81}
        \DataTypeTok{dob_conflict =} \KeywordTok{case_when}\NormalTok{(     }
\NormalTok{                      (}\StringTok{`}\DataTypeTok{Q1-3_A~M_1979}\StringTok{`} \OperatorTok{!=}\StringTok{ `}\DataTypeTok{Q1-3_A~M_1981}\StringTok{`}\NormalTok{) }\OperatorTok{&}\StringTok{ }\OperatorTok{!}\KeywordTok{is.na}\NormalTok{(}\StringTok{`}\DataTypeTok{Q1-3_A~M_1981}\StringTok{`}\NormalTok{)}
                      \OperatorTok{~}\StringTok{ }\OtherTok{TRUE}\NormalTok{,}
\NormalTok{                      (}\StringTok{`}\DataTypeTok{Q1-3_A~Y_1979}\StringTok{`} \OperatorTok{!=}\StringTok{ `}\DataTypeTok{Q1-3_A~Y_1981}\StringTok{`}\NormalTok{) }\OperatorTok{&}\StringTok{ }\OperatorTok{!}\KeywordTok{is.na}\NormalTok{(}\StringTok{`}\DataTypeTok{Q1-3_A~Y_1981}\StringTok{`}\NormalTok{)}
                      \OperatorTok{~}\StringTok{ }\OtherTok{TRUE}\NormalTok{,}
\NormalTok{                      (}\StringTok{`}\DataTypeTok{Q1-3_A~Y_1979}\StringTok{`} \OperatorTok{==}\StringTok{ `}\DataTypeTok{Q1-3_A~Y_1981}\StringTok{`}\NormalTok{) }\OperatorTok{&}\StringTok{ }
\StringTok{                      }\NormalTok{(}\StringTok{`}\DataTypeTok{Q1-3_A~M_1979}\StringTok{`} \OperatorTok{==}\StringTok{ `}\DataTypeTok{Q1-3_A~M_1981}\StringTok{`}\NormalTok{) }\OperatorTok{~}\StringTok{ }\OtherTok{FALSE}\NormalTok{,}
                      \KeywordTok{is.na}\NormalTok{(}\StringTok{`}\DataTypeTok{Q1-3_A~M_1981}\StringTok{`}\NormalTok{) }\OperatorTok{|}\StringTok{ }\KeywordTok{is.na}\NormalTok{(}\StringTok{`}\DataTypeTok{Q1-3_A~Y_1981}\StringTok{`}\NormalTok{) }\OperatorTok{~}\StringTok{ }\OtherTok{FALSE}\NormalTok{)) }\OperatorTok{%>%}
\StringTok{  }\NormalTok{dplyr}\OperatorTok{::}\KeywordTok{select}\NormalTok{(CASEID_}\DecValTok{1979}\NormalTok{,}
\NormalTok{         dob_month,}
\NormalTok{         dob_year,}
\NormalTok{         dob_conflict)}

@@

\begin{Shaded}
\begin{Highlighting}[]
\CommentTok{## tidy the gender and race variables}
\NormalTok{demog_tidy <-}\StringTok{ }\NormalTok{categories_qnames }\OperatorTok{%>%}
\StringTok{  }\NormalTok{dplyr}\OperatorTok{::}\KeywordTok{select}\NormalTok{(CASEID_}\DecValTok{1979}\NormalTok{,}
\NormalTok{         SAMPLE_RACE_78SCRN,}
\NormalTok{         SAMPLE_SEX_}\DecValTok{1979}\NormalTok{) }\OperatorTok{%>%}
\StringTok{  }\KeywordTok{rename}\NormalTok{(}\DataTypeTok{gender =}\NormalTok{ SAMPLE_SEX_}\DecValTok{1979}\NormalTok{,}
         \DataTypeTok{race =}\NormalTok{ SAMPLE_RACE_78SCRN)}
\end{Highlighting}
\end{Shaded}

@@

\begin{Shaded}
\begin{Highlighting}[]
\CommentTok{# make a list for years where we used the "QES-52A"}
\NormalTok{year_A <-}\StringTok{ }\KeywordTok{c}\NormalTok{(}\DecValTok{1979}\OperatorTok{:}\DecValTok{1987}\NormalTok{, }\DecValTok{1993}\NormalTok{)}
\CommentTok{#function to get the hour of work}
\NormalTok{get_hour <-}\StringTok{ }\ControlFlowTok{function}\NormalTok{(year)\{}
  \ControlFlowTok{if}\NormalTok{(year }\OperatorTok{%in%}\StringTok{ }\NormalTok{year_A)\{}
\NormalTok{   temp <-}\StringTok{ }\NormalTok{new_data_qnames }\OperatorTok{%>%}
\StringTok{    }\NormalTok{dplyr}\OperatorTok{::}\KeywordTok{select}\NormalTok{(CASEID_}\DecValTok{1979}\NormalTok{,}
            \KeywordTok{starts_with}\NormalTok{(}\StringTok{"QES-52A"}\NormalTok{) }\OperatorTok{&}
\StringTok{              }\KeywordTok{ends_with}\NormalTok{(}\KeywordTok{as.character}\NormalTok{(year)))\} }
  \ControlFlowTok{else}\NormalTok{\{}
\NormalTok{    temp <-}\StringTok{ }\NormalTok{new_data_qnames }\OperatorTok{%>%}
\StringTok{    }\NormalTok{dplyr}\OperatorTok{::}\KeywordTok{select}\NormalTok{(CASEID_}\DecValTok{1979}\NormalTok{,}
            \KeywordTok{starts_with}\NormalTok{(}\StringTok{"QES-52D"}\NormalTok{) }\OperatorTok{&}
\StringTok{              }\KeywordTok{ends_with}\NormalTok{(}\KeywordTok{as.character}\NormalTok{(year)))\} }
\NormalTok{    temp }\OperatorTok{%>%}\StringTok{ }
\StringTok{      }\KeywordTok{pivot_longer}\NormalTok{(}\OperatorTok{!}\NormalTok{CASEID_}\DecValTok{1979}\NormalTok{,}
                 \DataTypeTok{names_to =} \StringTok{"job"}\NormalTok{,}
                 \DataTypeTok{values_to =} \StringTok{"hours_work"}\NormalTok{) }\OperatorTok{%>%}
\StringTok{      }\KeywordTok{separate}\NormalTok{(}\StringTok{"job"}\NormalTok{, }\KeywordTok{c}\NormalTok{(}\StringTok{"job"}\NormalTok{, }\StringTok{"year"}\NormalTok{), }\DataTypeTok{sep =} \DecValTok{-4}\NormalTok{) }\OperatorTok{%>%}
\StringTok{      }\KeywordTok{mutate}\NormalTok{(}\DataTypeTok{job =} \KeywordTok{paste0}\NormalTok{(}\StringTok{"job_"}\NormalTok{, }\KeywordTok{substr}\NormalTok{(job, }\DecValTok{9}\NormalTok{, }\DecValTok{10}\NormalTok{))) }\OperatorTok{%>%}
\StringTok{      }\KeywordTok{rename}\NormalTok{(}\DataTypeTok{id =}\NormalTok{ CASEID_}\DecValTok{1979}\NormalTok{)}
\NormalTok{\}}

\CommentTok{# list to save the iteration result}
\NormalTok{hours <-}\StringTok{ }\KeywordTok{list}\NormalTok{()}
\CommentTok{# getting the hours of work of all observations}
\ControlFlowTok{for}\NormalTok{(ayear }\ControlFlowTok{in} \KeywordTok{c}\NormalTok{(}\DecValTok{1979}\OperatorTok{:}\DecValTok{1994}\NormalTok{, }\DecValTok{1996}\NormalTok{, }\DecValTok{1998}\NormalTok{, }\DecValTok{2000}\NormalTok{, }\DecValTok{2002}\NormalTok{, }\DecValTok{2004}\NormalTok{, }\DecValTok{2006}\NormalTok{, }\DecValTok{2008}\NormalTok{, }\DecValTok{2010}\NormalTok{, }
               \DecValTok{2012}\NormalTok{, }\DecValTok{2014}\NormalTok{, }\DecValTok{2016}\NormalTok{, }\DecValTok{2018}\NormalTok{)) \{}
\NormalTok{   hours[[ayear]] <-}\StringTok{ }\KeywordTok{get_hour}\NormalTok{(ayear)}
\NormalTok{\}}
\CommentTok{# unlist the hours of work}
\NormalTok{hours_all <-}\StringTok{ }\KeywordTok{bind_rows}\NormalTok{(}\OperatorTok{!!!}\NormalTok{hours)}
\end{Highlighting}
\end{Shaded}

@@

According to Huebner, Vach, and Cessie (2016), Initial Data Analysis (IDA) is the step of inspecting and screening the data after being collected to ensure that the data is clean, valid, and ready to be deployed in the later formal statistical analysis. Moreover, Chatfield (1985) argued that the two main objectives of IDA is data description, which is to assess the structure and the quality of the data; and model formulation without any formal statistical inference.

@@

\caption{\label{tab:age-table}Age Distribution of the NLSY79 samples}
\centering
\begin{tabular}[t]{r|r}
\hline
Age & Number of Sample\\
\hline
\rowcolor{gray!6}  15 & 1265\\
\hline
16 & 1550\\
\hline
\rowcolor{gray!6}  17 & 1600\\
\hline
18 & 1530\\
\hline
\rowcolor{gray!6}  19 & 1662\\
\hline
20 & 1722\\
\hline
\rowcolor{gray!6}  21 & 1677\\
\hline
22 & 1680\\
\hline
\end{tabular}
\end{table}

@@

\caption{\label{tab:gender-race-table}Gender and Race Distribution of the NLSY79 Samples}
\centering
\begin{tabular}[t]{l|l|l|l|l}
\hline
\multicolumn{1}{c|}{ } & \multicolumn{3}{c|}{Race} & \multicolumn{1}{c}{ } \\
\cline{2-4}
Gender & Hispanic & Black & Non-Black, Non-Hispanic & Total\\
\hline
\rowcolor{gray!6}  Male & 1000 (15.62\%) & 1613 (25.19\%) & 3790 (59.19\%) & 6403 (100.00\%)\\
\hline
Female & 1002 (15.95\%) & 1561 (24.84\%) & 3720 (59.21\%) & 6283 (100.00\%)\\
\hline
\rowcolor{gray!6}  Total & 2002 (15.78\%) & 3174 (25.02\%) & 7510 (59.20\%) & 12686 (100.00\%)\\
\hline
\end{tabular}
\end{table}

@@

>>>>

\title{The Journey from Wild to Text Book Data: A Case Study from the National Longitudinal Survey of Youth}

@@

% Pandoc citation processing
\newlength{\csllabelwidth}
\setlength{\csllabelwidth}{3em}
\newlength{\cslhangindent}
\setlength{\cslhangindent}{1.5em}
% for Pandoc 2.8 to 2.10.1
\newenvironment{cslreferences}%
  {}%
  {\par}
% For Pandoc 2.11+
\newenvironment{CSLReferences}[2] % #1 hanging-ident, #2 entry spacing
 {% don't indent paragraphs
  \setlength{\parindent}{0pt}
  % turn on hanging indent if param 1 is 1
  \ifodd #1 \everypar{\setlength{\hangindent}{\cslhangindent}}\ignorespaces\fi
  % set entry spacing
  \ifnum #2 > 0
  \setlength{\parskip}{#2\baselineskip}
  \fi
 }%
 {}
\usepackage{calc} % for calculating minipage widths
\newcommand{\CSLBlock}[1]{#1\hfill\break}
\newcommand{\CSLLeftMargin}[1]{\parbox[t]{\csllabelwidth}{#1}}
\newcommand{\CSLRightInline}[1]{\parbox[t]{\linewidth - \csllabelwidth}{#1}\break}
\newcommand{\CSLIndent}[1]{\hspace{\cslhangindent}#1}

@@

``Open data'' is data that are freely accessible, modifiable, and shareable by anyone for any purpose (Open Knowledge Foundation 2021). This type of data can be useful as example data in statistical text books and for research purposes. However, open data are often, what we might call ``wild data,'' because it requires substantial cleaning and tidying to tame it into text book shape. Huebner Marianne, Vach, and Cessie (2016) emphasize that making the data cleaning process accountable and transparent is imperative. Documenting the data cleaning is essential for the integrity of downstream statistical analyses and model building (M. Huebner et al. 2020).

@@

Despite its importance, this IDA stage is often undervalued and neglected (Chatfield 1985). There are few research papers that document the data cleaning (Wickham 2014). Furthermore, the decisions made in this stage often go unreported in the sense that IDA is often performed in an unplanned and unstructured way, and is only shared among restricted parties (M. Huebner et al. 2020).

@@

\caption{\label{tab:untidy-data}The Untidy Form of the NLSY79 Raw Data}
\centering
\begin{tabular}[t]{r|r|r|r|r|r|r}
\hline
CASEID\_1979 & HRP1\_1979 & HRP2\_1979 & HRP3\_1979 & HRP4\_1979 & HRP5\_1979 & HRP1\_1980\\
\hline
\cellcolor{gray!6}{1} & \cellcolor{gray!6}{328} & \cellcolor{gray!6}{NA} & \cellcolor{gray!6}{NA} & \cellcolor{gray!6}{NA} & \cellcolor{gray!6}{NA} & \cellcolor{gray!6}{NA}\\
\hline
2 & 385 & NA & NA & NA & NA & 457\\
\hline
\cellcolor{gray!6}{3} & \cellcolor{gray!6}{365} & \cellcolor{gray!6}{NA} & \cellcolor{gray!6}{275} & \cellcolor{gray!6}{NA} & \cellcolor{gray!6}{NA} & \cellcolor{gray!6}{397}\\
\hline
4 & NA & NA & NA & NA & NA & NA\\
\hline
\cellcolor{gray!6}{5} & \cellcolor{gray!6}{310} & \cellcolor{gray!6}{375} & \cellcolor{gray!6}{NA} & \cellcolor{gray!6}{NA} & \cellcolor{gray!6}{NA} & \cellcolor{gray!6}{333}\\
\hline
6 & NA & NA & NA & 250 & NA & 275\\
\hline
\end{tabular}
\end{table}

Consequently, the data should be tidied and wrangled first to extract the demographic and employment variables that we want to put in the final data set. We mainly used \texttt{tidyr} (Wickham 2020b), \texttt{dplyr} (Wickham et al. 2020), and \texttt{stringr} (Wickham 2019) to do this job.

@@

\DocumentationTok{\#\# tidy the date of birth data}
\NormalTok{dob\_tidy }\OtherTok{\textless{}{-}}\NormalTok{ new\_data\_qnames }\SpecialCharTok{\%\textgreater{}\%}
\NormalTok{  dplyr}\SpecialCharTok{::}\FunctionTok{select}\NormalTok{(CASEID\_1979,}
         \FunctionTok{starts\_with}\NormalTok{(}\StringTok{"Q1{-}3\_A\textasciitilde{}"}\NormalTok{)) }\SpecialCharTok{\%\textgreater{}\%}
  \FunctionTok{mutate}\NormalTok{(}\AttributeTok{dob\_year =} \FunctionTok{case\_when}\NormalTok{(}
                    \CommentTok{\# if the years recorded in both sets match, take 79 data}
                    \StringTok{\textasciigrave{}}\AttributeTok{Q1{-}3\_A\textasciitilde{}Y\_1979}\StringTok{\textasciigrave{}} \SpecialCharTok{==} \StringTok{\textasciigrave{}}\AttributeTok{Q1{-}3\_A\textasciitilde{}Y\_1981}\StringTok{\textasciigrave{}} \SpecialCharTok{\textasciitilde{}} \StringTok{\textasciigrave{}}\AttributeTok{Q1{-}3\_A\textasciitilde{}Y\_1979}\StringTok{\textasciigrave{}}\NormalTok{,}
                    \CommentTok{\# if the year in the 81 set is missing, take the 79 data}
                    \FunctionTok{is.na}\NormalTok{(}\StringTok{\textasciigrave{}}\AttributeTok{Q1{-}3\_A\textasciitilde{}Y\_1981}\StringTok{\textasciigrave{}}\NormalTok{) }\SpecialCharTok{\textasciitilde{}} \StringTok{\textasciigrave{}}\AttributeTok{Q1{-}3\_A\textasciitilde{}Y\_1979}\StringTok{\textasciigrave{}}\NormalTok{,}
                    \CommentTok{\# if the sets don\textquotesingle{}t match for dob year, take the 79 data}
                    \StringTok{\textasciigrave{}}\AttributeTok{Q1{-}3\_A\textasciitilde{}Y\_1979}\StringTok{\textasciigrave{}} \SpecialCharTok{!=} \StringTok{\textasciigrave{}}\AttributeTok{Q1{-}3\_A\textasciitilde{}Y\_1981}\StringTok{\textasciigrave{}} \SpecialCharTok{\textasciitilde{}} \StringTok{\textasciigrave{}}\AttributeTok{Q1{-}3\_A\textasciitilde{}Y\_1979}\StringTok{\textasciigrave{}}\NormalTok{),}
        \AttributeTok{dob\_month =} \FunctionTok{case\_when}\NormalTok{(}
                    \CommentTok{\# if months recorded in both sets match, take 79 data}
                    \StringTok{\textasciigrave{}}\AttributeTok{Q1{-}3\_A\textasciitilde{}M\_1979}\StringTok{\textasciigrave{}} \SpecialCharTok{==} \StringTok{\textasciigrave{}}\AttributeTok{Q1{-}3\_A\textasciitilde{}M\_1981}\StringTok{\textasciigrave{}} \SpecialCharTok{\textasciitilde{}} \StringTok{\textasciigrave{}}\AttributeTok{Q1{-}3\_A\textasciitilde{}M\_1979}\StringTok{\textasciigrave{}}\NormalTok{,}
                    \CommentTok{\# if month in 81 set is missing, take the 79 data}
                    \FunctionTok{is.na}\NormalTok{(}\StringTok{\textasciigrave{}}\AttributeTok{Q1{-}3\_A\textasciitilde{}M\_1981}\StringTok{\textasciigrave{}}\NormalTok{) }\SpecialCharTok{\textasciitilde{}} \StringTok{\textasciigrave{}}\AttributeTok{Q1{-}3\_A\textasciitilde{}M\_1979}\StringTok{\textasciigrave{}}\NormalTok{,}
                    \CommentTok{\# if sets don\textquotesingle{}t match for dob month, take the 79 data}
                    \StringTok{\textasciigrave{}}\AttributeTok{Q1{-}3\_A\textasciitilde{}M\_1979}\StringTok{\textasciigrave{}} \SpecialCharTok{!=} \StringTok{\textasciigrave{}}\AttributeTok{Q1{-}3\_A\textasciitilde{}M\_1981}\StringTok{\textasciigrave{}} \SpecialCharTok{\textasciitilde{}} \StringTok{\textasciigrave{}}\AttributeTok{Q1{-}3\_A\textasciitilde{}M\_1979}\StringTok{\textasciigrave{}}\NormalTok{),}
        \CommentTok{\# flag if there is a conflict between dob recorded in 79 and 81}
        \AttributeTok{dob\_conflict =} \FunctionTok{case\_when}\NormalTok{(     }
\NormalTok{                      (}\StringTok{\textasciigrave{}}\AttributeTok{Q1{-}3\_A\textasciitilde{}M\_1979}\StringTok{\textasciigrave{}} \SpecialCharTok{!=} \StringTok{\textasciigrave{}}\AttributeTok{Q1{-}3\_A\textasciitilde{}M\_1981}\StringTok{\textasciigrave{}}\NormalTok{) }\SpecialCharTok{\&} \SpecialCharTok{!}\FunctionTok{is.na}\NormalTok{(}\StringTok{\textasciigrave{}}\AttributeTok{Q1{-}3\_A\textasciitilde{}M\_1981}\StringTok{\textasciigrave{}}\NormalTok{)}
                      \SpecialCharTok{\textasciitilde{}} \ConstantTok{TRUE}\NormalTok{,}
\NormalTok{                      (}\StringTok{\textasciigrave{}}\AttributeTok{Q1{-}3\_A\textasciitilde{}Y\_1979}\StringTok{\textasciigrave{}} \SpecialCharTok{!=} \StringTok{\textasciigrave{}}\AttributeTok{Q1{-}3\_A\textasciitilde{}Y\_1981}\StringTok{\textasciigrave{}}\NormalTok{) }\SpecialCharTok{\&} \SpecialCharTok{!}\FunctionTok{is.na}\NormalTok{(}\StringTok{\textasciigrave{}}\AttributeTok{Q1{-}3\_A\textasciitilde{}Y\_1981}\StringTok{\textasciigrave{}}\NormalTok{)}
                      \SpecialCharTok{\textasciitilde{}} \ConstantTok{TRUE}\NormalTok{,}
\NormalTok{                      (}\StringTok{\textasciigrave{}}\AttributeTok{Q1{-}3\_A\textasciitilde{}Y\_1979}\StringTok{\textasciigrave{}} \SpecialCharTok{==} \StringTok{\textasciigrave{}}\AttributeTok{Q1{-}3\_A\textasciitilde{}Y\_1981}\StringTok{\textasciigrave{}}\NormalTok{) }\SpecialCharTok{\&} 
\NormalTok{                      (}\StringTok{\textasciigrave{}}\AttributeTok{Q1{-}3\_A\textasciitilde{}M\_1979}\StringTok{\textasciigrave{}} \SpecialCharTok{==} \StringTok{\textasciigrave{}}\AttributeTok{Q1{-}3\_A\textasciitilde{}M\_1981}\StringTok{\textasciigrave{}}\NormalTok{) }\SpecialCharTok{\textasciitilde{}} \ConstantTok{FALSE}\NormalTok{,}
                      \FunctionTok{is.na}\NormalTok{(}\StringTok{\textasciigrave{}}\AttributeTok{Q1{-}3\_A\textasciitilde{}M\_1981}\StringTok{\textasciigrave{}}\NormalTok{) }\SpecialCharTok{|} \FunctionTok{is.na}\NormalTok{(}\StringTok{\textasciigrave{}}\AttributeTok{Q1{-}3\_A\textasciitilde{}Y\_1981}\StringTok{\textasciigrave{}}\NormalTok{) }\SpecialCharTok{\textasciitilde{}} \ConstantTok{FALSE}\NormalTok{)) }\SpecialCharTok{\%\textgreater{}\%}
\NormalTok{  dplyr}\SpecialCharTok{::}\FunctionTok{select}\NormalTok{(CASEID\_1979,}
\NormalTok{         dob\_month,}
\NormalTok{         dob\_year,}
\NormalTok{         dob\_conflict)}

@@

\begin{Shaded}
\begin{Highlighting}[]
\DocumentationTok{\#\# tidy the gender and race variables}
\NormalTok{demog\_tidy }\OtherTok{\textless{}{-}}\NormalTok{ categories\_qnames }\SpecialCharTok{\%\textgreater{}\%}
\NormalTok{  dplyr}\SpecialCharTok{::}\FunctionTok{select}\NormalTok{(CASEID\_1979,}
\NormalTok{         SAMPLE\_RACE\_78SCRN,}
\NormalTok{         SAMPLE\_SEX\_1979) }\SpecialCharTok{\%\textgreater{}\%}
  \FunctionTok{rename}\NormalTok{(}\AttributeTok{gender =}\NormalTok{ SAMPLE\_SEX\_1979,}
         \AttributeTok{race =}\NormalTok{ SAMPLE\_RACE\_78SCRN)}
\end{Highlighting}
\end{Shaded}

@@

\begin{Shaded}
\begin{Highlighting}[]
\CommentTok{\# make a list for years where we used the "QES{-}52A"}
\NormalTok{year\_A }\OtherTok{\textless{}{-}} \FunctionTok{c}\NormalTok{(}\DecValTok{1979}\SpecialCharTok{:}\DecValTok{1987}\NormalTok{, }\DecValTok{1993}\NormalTok{)}
\CommentTok{\#function to get the hour of work}
\NormalTok{get\_hour }\OtherTok{\textless{}{-}} \ControlFlowTok{function}\NormalTok{(year)\{}
  \ControlFlowTok{if}\NormalTok{(year }\SpecialCharTok{\%in\%}\NormalTok{ year\_A)\{}
\NormalTok{   temp }\OtherTok{\textless{}{-}}\NormalTok{ new\_data\_qnames }\SpecialCharTok{\%\textgreater{}\%}
\NormalTok{    dplyr}\SpecialCharTok{::}\FunctionTok{select}\NormalTok{(CASEID\_1979,}
            \FunctionTok{starts\_with}\NormalTok{(}\StringTok{"QES{-}52A"}\NormalTok{) }\SpecialCharTok{\&}
              \FunctionTok{ends\_with}\NormalTok{(}\FunctionTok{as.character}\NormalTok{(year)))\} }
  \ControlFlowTok{else}\NormalTok{\{}
\NormalTok{    temp }\OtherTok{\textless{}{-}}\NormalTok{ new\_data\_qnames }\SpecialCharTok{\%\textgreater{}\%}
\NormalTok{    dplyr}\SpecialCharTok{::}\FunctionTok{select}\NormalTok{(CASEID\_1979,}
            \FunctionTok{starts\_with}\NormalTok{(}\StringTok{"QES{-}52D"}\NormalTok{) }\SpecialCharTok{\&}
              \FunctionTok{ends\_with}\NormalTok{(}\FunctionTok{as.character}\NormalTok{(year)))\} }
\NormalTok{    temp }\SpecialCharTok{\%\textgreater{}\%} 
      \FunctionTok{pivot\_longer}\NormalTok{(}\SpecialCharTok{!}\NormalTok{CASEID\_1979,}
                 \AttributeTok{names\_to =} \StringTok{"job"}\NormalTok{,}
                 \AttributeTok{values\_to =} \StringTok{"hours\_work"}\NormalTok{) }\SpecialCharTok{\%\textgreater{}\%}
      \FunctionTok{separate}\NormalTok{(}\StringTok{"job"}\NormalTok{, }\FunctionTok{c}\NormalTok{(}\StringTok{"job"}\NormalTok{, }\StringTok{"year"}\NormalTok{), }\AttributeTok{sep =} \SpecialCharTok{{-}}\DecValTok{4}\NormalTok{) }\SpecialCharTok{\%\textgreater{}\%}
      \FunctionTok{mutate}\NormalTok{(}\AttributeTok{job =} \FunctionTok{paste0}\NormalTok{(}\StringTok{"job\_"}\NormalTok{, }\FunctionTok{substr}\NormalTok{(job, }\DecValTok{9}\NormalTok{, }\DecValTok{10}\NormalTok{))) }\SpecialCharTok{\%\textgreater{}\%}
      \FunctionTok{rename}\NormalTok{(}\AttributeTok{id =}\NormalTok{ CASEID\_1979)}
\NormalTok{\}}

\CommentTok{\# list to save the iteration result}
\NormalTok{hours }\OtherTok{\textless{}{-}} \FunctionTok{list}\NormalTok{()}
\CommentTok{\# getting the hours of work of all observations}
\ControlFlowTok{for}\NormalTok{(ayear }\ControlFlowTok{in} \FunctionTok{c}\NormalTok{(}\DecValTok{1979}\SpecialCharTok{:}\DecValTok{1994}\NormalTok{, }\DecValTok{1996}\NormalTok{, }\DecValTok{1998}\NormalTok{, }\DecValTok{2000}\NormalTok{, }\DecValTok{2002}\NormalTok{, }\DecValTok{2004}\NormalTok{, }\DecValTok{2006}\NormalTok{, }\DecValTok{2008}\NormalTok{, }\DecValTok{2010}\NormalTok{, }
               \DecValTok{2012}\NormalTok{, }\DecValTok{2014}\NormalTok{, }\DecValTok{2016}\NormalTok{, }\DecValTok{2018}\NormalTok{)) \{}
\NormalTok{   hours[[ayear]] }\OtherTok{\textless{}{-}} \FunctionTok{get\_hour}\NormalTok{(ayear)}
\NormalTok{\}}
\CommentTok{\# unlist the hours of work}
\NormalTok{hours\_all }\OtherTok{\textless{}{-}} \FunctionTok{bind\_rows}\NormalTok{(}\SpecialCharTok{!!!}\NormalTok{hours)}
\end{Highlighting}
\end{Shaded}

@@

According to Huebner Marianne, Vach, and Cessie (2016), Initial Data Analysis (IDA) is the step of inspecting and screening the data after being collected to ensure that the data is clean, valid, and ready to be deployed in the later formal statistical analysis. Moreover, Chatfield (1985) argued that the two main objectives of IDA is data description, which is to assess the structure and the quality of the data; and model formulation without any formal statistical inference.

@@

\caption{\label{tab:age-table}Age Distribution of the NLSY79 samples}
\centering
\begin{tabular}[t]{r|r}
\hline
Age & Number of Sample\\
\hline
\cellcolor{gray!6}{15} & \cellcolor{gray!6}{1265}\\
\hline
16 & 1550\\
\hline
\cellcolor{gray!6}{17} & \cellcolor{gray!6}{1600}\\
\hline
18 & 1530\\
\hline
\cellcolor{gray!6}{19} & \cellcolor{gray!6}{1662}\\
\hline
20 & 1722\\
\hline
\cellcolor{gray!6}{21} & \cellcolor{gray!6}{1677}\\
\hline
22 & 1680\\
\hline
\end{tabular}
\end{table}

@@

\caption{\label{tab:gender-race-table}Gender and Race Distribution of the NLSY79 Samples}
\centering
\begin{tabular}[t]{l|l|l|l|l}
\hline
\multicolumn{1}{c|}{ } & \multicolumn{3}{c|}{Race} & \multicolumn{1}{c}{ } \\
\cline{2-4}
Gender & Hispanic & Black & Non-Black, Non-Hispanic & Total\\
\hline
\cellcolor{gray!6}{Male} & \cellcolor{gray!6}{1000 (15.62\%)} & \cellcolor{gray!6}{1613 (25.19\%)} & \cellcolor{gray!6}{3790 (59.19\%)} & \cellcolor{gray!6}{6403 (100.00\%)}\\
\hline
Female & 1002 (15.95\%) & 1561 (24.84\%) & 3720 (59.21\%) & 6283 (100.00\%)\\
\hline
\cellcolor{gray!6}{Total} & \cellcolor{gray!6}{2002 (15.78\%)} & \cellcolor{gray!6}{3174 (25.02\%)} & \cellcolor{gray!6}{7510 (59.20\%)} & \cellcolor{gray!6}{12686 (100.00\%)}\\
\hline
\end{tabular}
\end{table}

@@

\begin{Shaded}
\begin{Highlighting}[]
\FunctionTok{ggplot}\NormalTok{(wages\_slope, }\FunctionTok{aes}\NormalTok{(}\AttributeTok{x =}\NormalTok{ .slope\_year0)) }\SpecialCharTok{+}
  \FunctionTok{geom\_histogram}\NormalTok{()}
\end{Highlighting}
\end{Shaded}

\includegraphics{figures/unnamed-chunk-2-1.pdf}
